---
title: "Prediction and Bayesian Inference: Coding Exercise"
output: pdf_document
---

\renewcommand{\vec}[1]{\mathbf{#1}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 3, fig.width = 5, fig.align = 'center')
library(tidyverse) 
library(gridExtra)
library(rstanarm)
library(arm)
set.seed(09272020)
```



#### Contrasts and Functions of Parameters

Bayesian inference also allows easy computation of contrasts and more generally functions of parameters.

\vfill

Consider the contrast (predicted mean difference) between a weekend day with `max_tmp` = 20 and a weekday with `max_tmp` = 30. This would be a difficult problem to solve analytically, but it is straightforward using simulation.

\vfill

Note that a contrast of this sort (or any time), could also be calculated from a classical perspective, but, strictly speaking, the simulation approach is not permitted. Rather an analytical calculation, likely with a normal approximation (delta method?) would be necessary.

\vfill
1. Fit the model.
```{r}
beer <- read_csv('http://math.montana.edu/ahoegh/Data/Brazil_cerveja.csv')

stan_fit2 <- stan_glm(consumed ~ weekend + max_tmp, data = beer, refresh = 0)
print(stan_fit2)
```

\vfill

2. Extract the simulations.

\vfill
```{r}
contrast_sims <- as.matrix(stan_fit2)
```

\vfill

3. Compare differences
\vfill
```{r}
diff_consumed <- as.numeric(contrast_sims %*% matrix(c(1,1,20,0)) - 
  contrast_sims %*% matrix(c(1,0,30,0))) 
```
\vfill

4. Calculate interval and plot difference

```{r, echo = F}
quantile(diff_consumed, probs = c(.025, .975))
tibble(diff = diff_consumed) %>% ggplot(aes(x = diff)) + geom_histogram(bins = 100) + xlim(NA, 0) + 
  ggtitle('Distribution for predicted mean consumption \n 20 degree Weekend - 30 degree Weekday') + theme_minimal() +
  xlab('Difference (L)') + ylab('') + 
  theme(axis.ticks.y = element_blank(), 
        axis.text.y = element_blank(),
        axis.line.y = element_blank())   
```


### Prediction
1. Predict the expected (or average) consumption for max temperature of 25, weekend = 1, and precip = 0. 
\vfill

```{r}
pred_model <- stan_glm(consumed ~ max_tmp + weekend + precip, data = beer, refresh = 0)
print(pred_model)
x_prime <- tibble(max_tmp = 25, weekend = 1, precip = 0)
predictions <- posterior_linpred(pred_model, newdata = x_prime) %>% as.numeric()
predictions %>% quantile( probs = c(.025, .975))
predictions
```

\vfill

2. Predict the consumption for a new day with a max temperature of 25, weekday = 1, and precip = 0. 
\vfill
```{r}
daily_pred <- posterior_predict(pred_model, newdata = x_prime) %>% as.numeric()
daily_pred %>% quantile( probs = c(.025, .975))
```

\vfill

3. Create a figure to show a comparison of the predicted distributions.
```{r}
library(ggridges)
tibble( values = c(predictions,daily_pred), `Prediction Type` = rep(c('average prediction','individual prediction'),
                                                                 each = length(predictions))) %>% 
  ggplot(aes(x = values, y = `Prediction Type`, fill = `Prediction Type`)) + geom_density_ridges2(scale = 1) + 
  scale_y_discrete(expand = c(0.01, .01,0.01,.7)) + ggtitle('Predicted Beer Consumption in Brazil \n for weekend with max temp of 25 and no precip') + 
  ylab('') + xlab('liters of beer') +
  theme_bw() + theme(legend.position = "none") +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) 
```

